Practical Parametricity for GADTs

### 2020-09-28 20-gadts/inf-rules.tex
sections 1 through 3 of inf-rules.tex and IEL part of supp.tex done

\[\begin{array}{c}
\AXC{$\Gamma; \Phi \vdash F$}
\AXC{$\Gamma; \Phi \vdash G$}
\AXC{$\Gamma; \Phi \,|\, \Delta, x : F \vdash t: G$} 
\TIC{$\Gamma; \emptyset
  \,|\, \Delta \vdash L_\Phi x.t : \Nat^\Phi \,F \,G$}
\DisplayProof
\\\\
\AXC{$\ol{\Gamma;\Phi,\ol\beta \vdash K}$}
\AXC{$\Gamma; \emptyset
  \,|\, \Delta \vdash t : \Nat^{\ol\psi} \,F \,G$}
\AXC{$\Gamma;\Phi \,|\, \Delta \vdash s: F[\overline{\psi :=_{\ol\beta} K}]$}
\TIC{$\Gamma;\Phi\,|\, \Delta \vdash t_{\ol K} s:
  G[\overline{\psi :=_{\ol\beta} K}]$}
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi, \ol{\phi} \vdash H$}
\AXC{$\ol{\Gamma; \Phi,\ol{\beta} \vdash F}$}
\AXC{$\ol{\Gamma; \Phi,\ol{\beta} \vdash
    G}$}
\TIC{$\Gamma; \emptyset
  ~|~\emptyset
  \vdash \map^{\ol{F},\ol{G}}_H :
  \Nat^\emptyset\;(\ol{\Nat^{\Phi,\ol{\beta}}\,F\,G})\;
  (\Nat^\Phi\,H[\ol{\phi :=_{\ol{\beta}} F}]\;H[\ol{\phi
      :=_{\ol{\beta}} G}])$} 
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi, \phi, \ol{\alpha} \vdash H$}
\UIC{$\Gamma; \emptyset  \,|\, \emptyset \vdash \tin_H :
  \Nat^{\Phi,\ol{\beta}} H[\phi :=_{\ol{\beta}} (\mu
    \phi.\lambda \ol{\alpha}.H)\ol{\beta}][\ol{\alpha := \beta}]\,(\mu
  \phi.\lambda \ol{\alpha}.H)\ol{\beta}$}
\DisplayProof
\\\\
\AXC{$\Gamma; \phi,\Phi,\ol{\alpha} \vdash H$}
\AXC{$\Gamma; \Phi, \ol{\beta} \vdash F$}
\BIC{$\Gamma; \emptyset  \,|\, \emptyset \vdash \fold^F_H :
  \Nat^\emptyset\; (\Nat^{\Phi,\ol{\beta}}\,H[\phi
    :=_{\ol{\beta}} F][\ol{\alpha := \beta}]\,F)\; (\Nat^{\Phi,\ol{\beta}}
    \,(\mu \phi.\lambda \ol{\alpha}.H)\ol{\beta}\,F)$}
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi,\ol{\alpha}\vdash F$}
\AXC{$\ol{\Gamma; \ol{\alpha} \vdash K}$}
%\AXC{$\ol{\Gamma;\Phi \vdash A} \hspace*{0.3in} \Gamma;\Phi~|~\Delta \vdash t : F[\ol{\alpha :=A}]$}
\BIC{$\Gamma;\emptyset~|~\Delta \vdash \int_{\ol K,F} :
\Nat^{\Phi,\ol\alpha}\,F\, (\Lan^{\ol\alpha}_{\ol K}\,  F)\ol{K}$}  
\DisplayProof
\\\\
\AXC{$\Gamma;\emptyset~|~\Delta \vdash \eta : \Nat^{\Phi,\ol\alpha}\,
  F\;G[\ol{\beta := K}]$}
%\AXC{$\ol{\Gamma;\Phi \vdash B}$}
%\AXC{$\Gamma;\Phi~|~\Delta \vdash t : (\Lan^{\ol\alpha}_{\ol K} F)\,\ol B $}
\UIC{$\Gamma;\emptyset~|~\Delta \vdash \partial^{G, \ol K}_F \eta :
\Nat^{\Phi, \ol\beta}\, (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta\; G$}
\DisplayProof
\end{array}\]
\end{dfn}

@@

In the rule for $L_{\ol{\alpha}}x.t$, the $L$ operator binds all
occurrences of the type variables in $\ol{\alpha}$ in the type of the
term variable $x$ and in the body $t$, as well as all occurrences of
$x$ in $t$. In the rule for $t_{\ol K} s$ there is one functorial
expression in $\ol K$ for every functorial variable in $\ol
\alpha$. In the rule for $\map^{\ol{F},\ol{G}}_H$ there is one
functorial expression $F$ and one functorial expression $G$ for each
functorial variable in $\ol\phi$. Moreover, for each $\phi^k$ in
$\ol\phi$ the number of functorial variables in $\ol\beta$ in the
judgments for its corresponding functorial expresssions $F$ and $G$ is
$k$. In the rules for $\tin_H$ and $\fold^F_H$, the functorial
variables in $\ol{\beta}$ are fresh with respect to $H$, and there is
one $\beta$ for every $\alpha$. (Recall from above that, in order for
the types of $\tin_H$ and $\fold^F_H$ to be well-formed, the length of
$\alpha$ must equal the arity of $\phi$.)  {\color{red} In the rule
  for $\int_{\ol K,H}$, ANYTHING NEEDED HERE?}

@@

\begin{dfn}\label{def:set-sem}
The {\em object interpretation} $\oCPOsem{\cdot} : \F \to [\oCPOenv,
  \oCPO]$ is defined by
\begin{align*}
  \oCPOsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \oCPOsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
  \oCPOsem{\Gamma; \emptyset
    \vdash \Nat^\Phi
    \,F\,G}\rho &= \{\eta : \lambda \ol{K}. \,\oCPOsem{\Gamma;
    \Phi \vdash
    F}\rho[\Phi := \ol K] 
      \Rightarrow \lambda \ol{K}.\,\oCPOsem{\Gamma; 
        \Phi \vdash G}\rho[\Phi := \ol K] \\ 
      &\hspace{0.3in}|~\forall \ol{K = (K^1, K^2, K^*) : \oCPORT_k}.\\
      &\hspace{0.4in}(\eta_{\overline{K^1}}, \eta_{\overline{K^2}})
      : \oCPORsem{\Gamma; \Phi \vdash F}\Eq_{\rho}[\Phi := \ol K]
      \rightarrow \oCPORsem{\Gamma; \Phi \vdash
        G}\Eq_{\rho}[\Phi := \ol K] \} \\
  \oCPOsem{\Gamma;\Phi \vdash \phi\ol{F}}\rho &=
  (\rho\phi)\,\ol{\oCPOsem{\Gamma;\Phi \vdash
    F}\rho}\\
  \oCPOsem{\Gamma;\Phi \vdash F+G}\rho &=
  \oCPOsem{\Gamma;\Phi \vdash F}\rho +
  \oCPOsem{\Gamma;\Phi \vdash G}\rho\\
  \oCPOsem{\Gamma;\Phi \vdash F \times G}\rho &=
  \oCPOsem{\Gamma;\Phi \vdash F}\rho \times
  \oCPOsem{\Gamma;\Phi \vdash G}\rho\\ 
  \oCPOsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho &= (\mu
    T^\oCPO_{H,\rho})\ol{\oCPOsem{\Gamma;\Phi,\ol\gamma \vdash G}\rho}\\
    \text{where } T^\oCPO_{H,\rho}\,F & = \lambda
  \ol{A}. \oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\rho[\phi :=  F][\ol{\alpha := A}]\\
  \text{and } T^\oCPO_{H,\rho}\,\eta &= \lambda
  \ol{A}. \oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\id_\rho[\phi := \eta][\ol{\alpha := \id_{A}}] \\
  \oCPOsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\rho &=
    \{ t : (\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
          K}\rho[\ol{\alpha := \_}]}} \oCPOsem{\Gamma; \Phi, \ol\alpha
      \vdash F}\rho[\ol{\alpha := \_}]) \ol{\oCPOsem{\Gamma; \Phi
        \vdash A}\rho}\\
    &\hspace{0.3in}|~(t,t) \in \oCPORsem{\Gamma;\Phi \vdash
      (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\Eq_\rho\}\\
\end{align*}
\end{dfn}
\new{WHERE DOES THIS ALL GO? Must explain $\Lan$s, say why they are
  functors, say they can be computed using colimits in $\set$ and
  $\rel$, define injections, etc. If $f : B \to B'$ in $\set$, then
  the functorial action of $\mathit{Lan}_{\ol K}\,F$ on $f$ is the
  unique morphism $(\mathit{Lan}_{\ol K}\,F)f$ such that
  \begin{equation}\label{def:functorial-action-lan}
(\mathit{Lan}_{\ol K}\,F)f \circ \iota_{X,g} = \iota'_{X, f \circ
      g}~~~MISSING~SOME~BARS
\end{equation}
Here, $\iota_{X,g}$ is the injection of $FX$ into $(\mathit{Lan}_{\ol
  K}\,F)\,B = \colim{X : \set_0,\, g : KX \to B}{FX}$ and
$\iota'_{X,g'}$ is the injection of $FX$ into $(\mathit{Lan}_{\ol
  K}\,F)\,B' = \colim{X : \set_0,\, g' : KX \to B'}{FX}$.}

@@

\begin{thm}[IEL]\label{thm:iel}
  If $\rho : \oCPOenv$ and $\Gamma; \Phi \vdash F$ then
  $\oCPORsem{\Gamma; \Phi \vdash F} \Eq_\rho = \Eq_{\oCPOsem{\Gamma;
      \Phi \vdash F}\rho}$.
\end{thm}
\begin{proof}
By induction on the structure of $F$. Only the $\Nat$, application,
fixpoint, and $\Lan$ cases are non-routine.

The latter two use
Equation~\ref{eq:eq-pres-eq}. The fixpoint case also uses the
observation that, for every $i < \omega$, the following intermediate
results can be proved by simultaneous induction with
Theorem~\ref{thm:iel}: {\color{red} CHECK THIS!} for any $H$, $\rho$,
$\ol A$, and any subformula $J$ of $H$, both $T^i_{H,\Eq_{\rho}} K_0\,
\ol{\Eq_A} = (\Eq_{(T^\set_{H,\rho})^i K_0})^* \ol{\Eq_A}$\; and\;
\[\begin{array}{ll}
&  \oCPORsem{\Gamma; \Phi, \phi, \ol{\alpha} \vdash J} \Eq_{\rho} [\phi
  := T^i_{H,\Eq_{\rho}} K_0] \overline{[\alpha := \Eq_A]}\\
=  &
\oCPORsem{\Gamma; \Phi, \phi,
  \ol{\alpha} \vdash J} \Eq_{\rho} [\phi := \Eq_{(T^\oCPO_{H,\rho})^i K_0}]
\overline{[\alpha := \Eq_A]}
\end{array}\]
hold. With these results in hand, the proof follows easily
{\color{red} for all but $\Lan$-types}. The entire proof is given in
detail in the accompanying anonymous supplementary material. As noted
there, if functorial variables of arity greater than $0$ were allowed
to appear in the bodies of $\mu$-types, then the IEL would fail.

{\color{red} For $\Lan$-types,....}

@@

\vspace*{1in}

\hrule

\new{
Some questions/issues:
\begin{itemize}
\item Can we write zipBush and appendBush with $\partial$ and $\int$?
  We could already represent the uncurried type of appendBush
  (although not its curried type), but couldn't recurse over both
  input bushes because folds take natural transformations as inputs.
\item More generally, how do we compute with $\partial$ and $\int?$
  Can we use the colimit formulation of Lans (see Lemma 6.3.7
  of~\cite{rie16}) to get a handle on this?
\item What is the connection between exponentials and natural
  transformations? (Should we assume only small objects are
  exponentiable?) Do we want the former or the latter for
  computational purposes? (I suspect the latter.)

  [From nlab:
  In a functor category $D^C$ , a natural transformation $\alpha : F \to
  G$ is exponentiable if (though probably not ``only if'') it is
  cartesian and each component $\alpha_c : F c \to G c$ is
  exponentiable in $D$.  Given $H \to F$ we define $(\Pi_\alpha H) c =
  \Pi_{\alpha_c}(Hc)$; then for $u : c \to c'$ to obtain a map
  $\Pi_{\alpha_c} (Hc) \to \Pi_{\alpha_{c'}} (Hc')$ we need a map
  $\alpha^*_{c'}(\Pi_{\alpha_c}(Hc)) \to Hc'$. But since $\alpha$ is
  cartesian, $\alpha^*_{c'}(\Pi_{\alpha_c}(Hc)) \cong
  \alpha^*_{c}(\Pi_{\alpha_c}(Hc))$, so we have the counit
  $\alpha^*_{c}(\Pi_{\alpha_c}(Hc)) \to Hc$ that we can compose with
  $Hu$.]
\item After we understand what we can do with Lans and folds on GADTs
  we might want to try to extend calculus with term-level
  fixpoints. This would give a categorical analogue for GADTs
  of~\cite{pit98,pit00} for ADTs. Would it also more accurately
  reflect how GADTs are used in practice, or are functions over GADTs
  usually folds? Investigate applications in the literature and/or in
  implementations.
\item $\omega$CPO is a natural choice for modeling general
  recursion. We know $(Lan^\gamma_C \onet)D$ is $C \to D$ for any
  closed type $C$. (Also for select classes of open types?) So can
  model $\Nat \to \gamma$. But the functor $NX = \Nat \to X$ isn't
  $\omega$-cocontinuous. It also doesn't preserve
  $\omega$-presentable objects, i.e., countable $\omega$CPOs since
  $\Nat \to \Nat$ is not countable. So we cannot have a functor like
  $N$ as the subscript to $\Lan$ and expect the resulting $\Lan$ to be
  $\omega$-cocontinuous.
\item What functors can be subscripts to $\Lan$ and produce
  $\omega$-cocontinuous functors? We can use functors that preserve
  presentable objects by theorem in~\cite{jp19}, and possibly others
  as well. These include polynomial functors, ADTs and nested types
  seen as functors, certain (which?) GADTs seen as functors? How big
  can GADTs get? 
\end{itemize}
}

>>>>

\[\begin{array}{c}
\AXC{$\Gamma; \Phi \vdash F$}
\AXC{$\Gamma; \Phi \vdash G$}
\AXC{$\Gamma; \Phi \,|\, \Delta, x : F \vdash t: G$} 
\TIC{$\Gamma; \emptyset
  \,|\, \Delta \vdash L_\Phi x.t : \Nat^\Phi \,F \,G$}
\DisplayProof
\\\\
\AXC{$\ol{\Gamma;\Phi,\ol\beta \vdash K}$}
\AXC{$\Gamma; \emptyset
  \,|\, \Delta \vdash t : \Nat^{\ol\psi} \,F \,G$}
\AXC{$\Gamma;\Phi \,|\, \Delta \vdash s: F[\overline{\psi :=_{\ol\beta} K}]$}
\TIC{$\Gamma;\Phi\,|\, \Delta \vdash t_{\ol K} s:
  G[\overline{\psi :=_{\ol\beta} K}]$}
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi, \ol{\phi} \vdash H$}
\AXC{$\ol{\Gamma; \Phi,\ol{\beta} \vdash F}$}
\AXC{$\ol{\Gamma; \Phi,\ol{\beta} \vdash
    G}$}
\TIC{$\Gamma; \emptyset
  ~|~\emptyset
  \vdash \map^{\ol{F},\ol{G}}_H :
  \Nat^\emptyset\;(\ol{\Nat^{\Phi,\ol{\beta}}\,F\,G})\;
  (\Nat^\Phi\,H[\ol{\phi :=_{\ol{\beta}} F}]\;H[\ol{\phi
      :=_{\ol{\beta}} G}])$} 
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi, \phi, \ol{\alpha} \vdash H$}
\UIC{$\Gamma; \emptyset  \,|\, \emptyset \vdash \tin_H :
  \Nat^{\Phi,\ol{\beta}} H[\phi :=_{\ol{\beta}} (\mu
    \phi.\lambda \ol{\alpha}.H)\ol{\beta}][\ol{\alpha := \beta}]\,(\mu
  \phi.\lambda \ol{\alpha}.H)\ol{\beta}$}
\DisplayProof
\\\\
\AXC{$\Gamma; \phi,\Phi,\ol{\alpha} \vdash H$}
\AXC{$\Gamma; \Phi, \ol{\beta} \vdash F$}
\BIC{$\Gamma; \emptyset  \,|\, \emptyset \vdash \fold^F_H :
  \Nat^\emptyset\; (\Nat^{\Phi,\ol{\beta}}\,H[\phi
    :=_{\ol{\beta}} F][\ol{\alpha := \beta}]\,F)\; (\Nat^{\Phi,\ol{\beta}}
    \,(\mu \phi.\lambda \ol{\alpha}.H)\ol{\beta}\,F)$}
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi,\ol{\alpha}\vdash F$}
\AXC{$\ol{\Gamma; \ol{\alpha} \vdash K}$}
%\AXC{$\ol{\Gamma;\Phi \vdash A} \hspace*{0.3in} \Gamma;\Phi~|~\Delta \vdash t : F[\ol{\alpha :=A}]$}
\BIC{$\Gamma;\emptyset~|~\emptyset \vdash \int_{\ol K,F} :
\Nat^{\Phi,\ol\alpha}\,F\, (\Lan^{\ol\alpha}_{\ol K}\,  F)\ol{K}$}  
\DisplayProof
\\\\
\AXC{$\Gamma;\emptyset~|~\Delta \vdash \eta : \Nat^{\Phi,\ol\alpha}\,
  F\;G[\ol{\beta := K}]$}
%\AXC{$\ol{\Gamma;\Phi \vdash B}$}
%\AXC{$\Gamma;\Phi~|~\Delta \vdash t : (\Lan^{\ol\alpha}_{\ol K} F)\,\ol B $}
\UIC{$\Gamma;\emptyset~|~\Delta \vdash \partial^{G, \ol K}_F \eta :
\Nat^{\Phi, \ol\beta}\, (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta\; G$}
\DisplayProof
\end{array}\]
\end{dfn}

@@

In the rule for $L_{\ol{\alpha}}x.t$, the $L$ operator binds all
occurrences of the type variables in $\ol{\alpha}$ in the type of the
term variable $x$ and in the body $t$, as well as all occurrences of
$x$ in $t$. In the rule for $t_{\ol K} s$ there is one functorial
expression in $\ol K$ for every functorial variable in $\ol
\alpha$. In the rule for $\map^{\ol{F},\ol{G}}_H$ there is one
functorial expression $F$ and one functorial expression $G$ for each
functorial variable in $\ol\phi$. Moreover, for each $\phi^k$ in
$\ol\phi$ the number of functorial variables in $\ol\beta$ in the
judgments for its corresponding functorial expresssions $F$ and $G$ is
$k$. In the rules for $\tin_H$ and $\fold^F_H$, the functorial
variables in $\ol{\beta}$ are fresh with respect to $H$, and there is
one $\beta$ for every $\alpha$. (Recall from above that, in order for
the types of $\tin_H$ and $\fold^F_H$ to be well-formed, the length of
$\alpha$ must equal the arity of $\phi$.)

@@

\begin{dfn}\label{def:set-sem}
The {\em object interpretation} $\oCPOsem{\cdot} : \F \to [\oCPOenv,
  \oCPO]$ is defined by
\begin{align*}
  \oCPOsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \oCPOsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
  \oCPOsem{\Gamma; \emptyset
    \vdash \Nat^\Phi
    \,F\,G}\rho &= \{\eta : \lambda \ol{K}. \,\oCPOsem{\Gamma;
    \Phi \vdash
    F}\rho[\Phi := \ol K] 
      \Rightarrow \lambda \ol{K}.\,\oCPOsem{\Gamma; 
        \Phi \vdash G}\rho[\Phi := \ol K] \\ 
      &\hspace{0.3in}|~\forall \ol{K = (K^1, K^2, K^*) : \oCPORT_k}.\\
      &\hspace{0.4in}(\eta_{\overline{K^1}}, \eta_{\overline{K^2}})
      : \oCPORsem{\Gamma; \Phi \vdash F}\Eq_{\rho}[\Phi := \ol K]
      \rightarrow \oCPORsem{\Gamma; \Phi \vdash
        G}\Eq_{\rho}[\Phi := \ol K] \} \\
  \oCPOsem{\Gamma;\Phi \vdash \phi\ol{F}}\rho &=
  (\rho\phi)\,\ol{\oCPOsem{\Gamma;\Phi \vdash
    F}\rho}\\
  \oCPOsem{\Gamma;\Phi \vdash F+G}\rho &=
  \oCPOsem{\Gamma;\Phi \vdash F}\rho +
  \oCPOsem{\Gamma;\Phi \vdash G}\rho\\
  \oCPOsem{\Gamma;\Phi \vdash F \times G}\rho &=
  \oCPOsem{\Gamma;\Phi \vdash F}\rho \times
  \oCPOsem{\Gamma;\Phi \vdash G}\rho\\ 
  \oCPOsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho &= (\mu
    T^\oCPO_{H,\rho})\ol{\oCPOsem{\Gamma;\Phi,\ol\gamma \vdash G}\rho}\\
    \text{where } T^\oCPO_{H,\rho}\,F & = \lambda
  \ol{A}. \oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\rho[\phi :=  F][\ol{\alpha := A}]\\
  \text{and } T^\oCPO_{H,\rho}\,\eta &= \lambda
  \ol{A}. \oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\id_\rho[\phi := \eta][\ol{\alpha := \id_{A}}] \\
  \oCPOsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\rho &=
    \{ t : (\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
          K}\rho[\ol{\alpha := \_}]}} \oCPOsem{\Gamma; \Phi, \ol\alpha
      \vdash F}\rho[\ol{\alpha := \_}]) \ol{\oCPOsem{\Gamma; \Phi
        \vdash A}\rho}\\
    &\hspace{0.3in}|~(t,t) \in \oCPORsem{\Gamma;\Phi \vdash
      (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\Eq_\rho\}\\
\end{align*}
\end{dfn}
\new{WHERE DOES THIS ALL GO? Must explain $\Lan$s, say why they are
  functors, say they can be computed using colimits in $\set$ and
  $\rel$, define injections, etc. If $f : B \to B'$ in $\set$, then
  the functorial action of $\mathit{Lan}_{\ol K}\,F$ on $f$ is the
  unique morphism $(\mathit{Lan}_{\ol K}\,F)f$ such that
  \begin{equation}\label{def:functorial-action-lan}
(\mathit{Lan}_{\ol K}\,F)f \circ \iota_{X,g} = \iota'_{X, f \circ
      g}~~~MISSING~SOME~BARS
\end{equation}
Here, $\iota_{X,g}$ is the injection of $FX$ into $(\mathit{Lan}_{\ol
  K}\,F)\,B = \colim{X : \set_0,\, g : KX \to B}{FX}$ and
$\iota'_{X,g'}$ is the injection of $FX$ into $(\mathit{Lan}_{\ol
  K}\,F)\,B' = \colim{X : \set_0,\, g' : KX \to B'}{FX}$.

**Say somewhere that recognizing the non-parallel nature of the Lan
interps, and yet still being able to define the set and rel interps so
that everything is intertwined in exactly the right way to make the
proofs (e.g., of the IEL) work, is what's hard here. This is where the
insight is required.**}

@@

\begin{thm}[IEL]\label{thm:iel}
  If $\rho : \oCPOenv$ and $\Gamma; \Phi \vdash F$ then
  $\oCPORsem{\Gamma; \Phi \vdash F} \Eq_\rho = \Eq_{\oCPOsem{\Gamma;
      \Phi \vdash F}\rho}$.
\end{thm}
\begin{proof}
By induction on the structure of $F$. Only the $\Nat$, application,
fixpoint, and $\Lan$ cases are non-routine. The application and
fixpoint cases use Equation~\ref{eq:eq-pres-eq}. The fixpoint case
$\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda \ol\alpha.H)\ol F$
also uses the observation that, for every $i < \omega$, the following
intermediate results can be proved by simultaneous induction with
Theorem~\ref{thm:iel}: for any $H$, $\rho$, $\ol A$, and any
subformula $J$ of $H$, $T^i_{H,\Eq_{\rho}} K_0\, \ol{\Eq_A} =
(\Eq_{(T^\set_{H,\rho})^i K_0})^* \ol{\Eq_A}$\; and\;
\[\begin{array}{ll}
&  \oCPORsem{\Gamma; \Phi, \phi, \ol{\alpha} \vdash J} \Eq_{\rho} [\phi
  := T^i_{H,\Eq_{\rho}} K_0] \overline{[\alpha := \Eq_A]}\\
=  &
\oCPORsem{\Gamma; \Phi, \phi,
  \ol{\alpha} \vdash J} \Eq_{\rho} [\phi := \Eq_{(T^\oCPO_{H,\rho})^i K_0}]
\overline{[\alpha := \Eq_A]}
\end{array}\]
The case of the latter equation for $J = (\Lan_{\ol K}^{\ol
  \beta}\,F)\ol B$ uses the facts that $\pi_1 (T^i_{H,\Eq_\rho} K_0) =
(T^\set_{H,\rho})^i K_0$ for all $i < \omega$, $\pi_2
(T^i_{H,\Eq_\rho} K_0) = (T^\set_{H,\rho})^i K_0$ for all $i <
\omega$, and that, without loss of generality, the variables in $\ol
\beta$ do not appear free in $H$. With these results in hand, the
proof of Theorem~\ref{thm:iel} follows. The simultaneous induction
required to make the various parts of the proof hang together
appropriately is somewhat delicate, so we give the proof in its
entirety in the {\color{blue} appendix}. As noted there, if functorial
variables of arity greater than $0$ were allowed in the bodies of
$\mu$-types then the IEL would fail.
\end{proof}

@@

\section{Interpreting Terms}
